Local Spectral Clustering of Density Upper Level Sets

### 2019-05-23 neurips/local_spectral_density_clustering_neurips.tex
Through Sect 3

\paragraph{Geometric conditions on density clusters.} At a high level, for PPR
to be successful, the underlying density cluster must be geometrically
well-conditioned.  At a minimum, we want to avoid sets that contain arbitrarily
thin bridges or spikes.  Hence, as in \cite{chaudhuri2010}, we consider a
thickened version of \smash{$\Cset \in \Cbb_f(\lambda)$} defined as 
$\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq \sigma}$, which 
we call the \emph{$\sigma$-expansion} of $\Cset$. Here 
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

@@

\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c_4 >
  0$ such that with probability tending to $1$ as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c_4 \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c_5 >
  0$ such that if 
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c_5 \frac{\lambda_{\sigma}^2r^d
      \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)}, 
  \end{equation}
  then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for
  $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}

@@

\begin{theorem}
	\label{thm: mixing_time_upper_bound}
	Fix $\lambda > 0$, and let $\Cset \in \Cbb_f(\lambda)$ satisfy Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding} for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, \rho, L > 0$. Then, for any $0 < r < \sigma/2\sqrt{d}$, with probability $1$
	\begin{equation}
	\label{eqn: mixing_time_upper_bound}
	\limsup_{n \to \infty}\tau_{\infty}(G_{n,r}[\Csig[\Xbf]]) \leq c_2 \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2} \log^2\left(\frac{1}{r}\right) + c_3
	\end{equation}
	for $c_2,c_3 > 0$ universal constants. 
\end{theorem}

To the best of our knowledge, Theorem \ref{thm: mixing_time_upper_bound} is the first bound, albeit asymptotic, on the mixing time of random walks over neighborhood graphs which is independent of $n$, the number of vertices.
\begin{remark}
	The embedding assumption \ref{asmp: embedding} and Lipschitz parameter $L$ play an important role in proving the upper bound of Theorem \ref{thm: mixing_time_upper_bound}. There is some interdependence between $L$ and other geometric parameters $\sigma$ and $\rho$, which might lead one to hope that \ref{asmp: embedding} is non-essential. However, it is not possible to eliminate this condition without incurring an additional factor of at least $(\rho/\sigma)^d$ in \eqref{eqn: mixing_time_upper_bound}, achieved, for instance, when $\Csig$ is a dumbbell-like set consisting of two balls of diameter $\rho$ linked by a cylinder of radius $\sigma$.  \citep{abbasi-yadkori2016, abbasi-yadkori2016a} develop theory regarding bi-Lipschitz deformations of convex sets, wherein it is observed that star-shaped sets as well as half-moon shapes of the type we consider in Section \ref{sec: experiments} both satisfy \ref{asmp: embedding} for reasonably small values of $L$.
\end{remark}

@@

\begin{figure}
	\centering
	\begin{adjustbox}{minipage=\linewidth}
		\begin{subfigure}{.33\linewidth}
			\includegraphics[width=\linewidth]{example1plots/sample2}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.33\linewidth}
			\includegraphics[width=\linewidth]{example1plots/sample1}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.33\linewidth}
			\includegraphics[width=\linewidth]{example1plots/sigma_normalized_cut_plot}
			\caption{}
		\end{subfigure}

		\begin{subfigure}{.33\linewidth}
			\includegraphics[width=\linewidth]{example1plots/sigma_mixing_time_plot}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.33\linewidth}
			\includegraphics[width=\linewidth]{example1plots/diameter_normalized_cut_plot}
			\caption{}
		\end{subfigure}
		\begin{subfigure}{.33\linewidth}
			\includegraphics[width=\linewidth]{example1plots/diameter_mixing_time_plot}
			\caption{}
		\end{subfigure}
		\caption{Samples, empirical results, and theoretical bounds for mixing time and normalized cut as diameter and thickness are varied. In (a) and (b), points in $\Cset$ are colored in red; points in $\Csig \setminus \Cset$ are colored in yellow; and remaining points in blue.}
		\label{fig:fig1}
	\end{adjustbox}
\end{figure}

>>>>

\paragraph{Geometric conditions on density clusters.} At a high level, for PPR
to be successful, the underlying density cluster must be geometrically
well-conditioned.  At a minimum, we want to avoid sets that contain arbitrarily
thin bridges or spikes.  Hence, as in \citet{chaudhuri2010}, we consider a
thickened version of \smash{$\Cset \in \Cbb_f(\lambda)$} defined as 
$\Csig := \set{x \in \Reals^d: \dist(x,\Cset) \leq \sigma}$, which 
we call the \emph{$\sigma$-expansion} of $\Cset$. Here 
\smash{$\dist(x,\Cset) := \inf_{y \in \Cset} \norm{y - x}$}.  We now list our
conditions on $\Csig$.

@@

\begin{theorem}
  \label{thm: misclassification_rate}
  Fix $\lambda > 0$, let \smash{$\Cset \in \Cbb_f(\lambda)$} be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: valid_region}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >
  0$ such that with probability tending to 1 as $n \to \infty$,  
  \begin{equation}
    \label{eqn: misclassification_rate_ub}
    \Delta(\Csig[\Xbf], \Cest) \leq c \kappa(\Cset)
    \frac{\Lambda_{\sigma}}{\lambda_{\min}}. 
  \end{equation}
\end{theorem}

@@

\begin{theorem}
  \label{thm: consistent_recovery_of_density_clusters}
  Fix $\lambda > 0$, let $\Cset \in \Cbb_f(\lambda)$ be a
  $\kappa$-well-conditioned density cluster (with respect to some $\theta$), and
  additionally assume $f$ satisfies \ref{asmp: C'_bounded_density}. If Algorithm
  \ref{alg: ppr} is well-initialized, there exists a universal constant $c >  0$
  such that if  
  \begin{equation}
    \label{eqn: kappa_ub}
    \kappa(\Cset) \leq c \frac{\lambda_{\sigma}^2r^d
      \nu_d}{\Lambda_{\sigma}\Pbb(\Csig)}, 
  \end{equation}
  then the output set $\Cest \subseteq \Xbf$ is a consistent estimator for
  $\Cset$, in the sense of Definition \ref{def: consistent_density_cluster_estimation}. 
\end{theorem}

@@

\begin{theorem}
  \label{thm: mixing_time_upper_bound}
  Fix $\lambda > 0$, and assume that \smash{$\Cset \in \Cbb_f(\lambda)$}
  satisfies Assumptions \ref{asmp: bounded_density} and \ref{asmp: embedding}
  for some $\sigma, \lambda_{\sigma}, \Lambda_{\sigma}, \rho, L > 0$. Then, for
  any \smash{$0 < r < \sigma/2\sqrt{d}$}, with probability 1,
  \begin{equation} 
    \label{eqn: mixing_time_upper_bound}
    \limsup_{n \to \infty}\tau_{\infty}(G_{n,r}[\Csig[\Xbf]]) \leq c_1
    \frac{\Lambda_{\sigma}^4 d^3 \rho^2 L^2}{\lambda_{\sigma}^4 r^2}
    \log^2\left(\frac{1}{r}\right) + c_2,
  \end{equation}
  for $c_1,c_2 > 0$ universal constants. 
\end{theorem}

To the best of our knowledge, Theorem \ref{thm: mixing_time_upper_bound} is the
first bound, albeit asymptotic, on the mixing time of random walks over
neighborhood graphs which is independent of $n$, the number of vertices.

\begin{remark}
  The embedding assumption \ref{asmp: embedding} and Lipschitz parameter $L$
  play an important role in proving the upper bound of Theorem \ref{thm:
    mixing_time_upper_bound}. There is some interdependence between $L$ and
  $\sigma,\rho$, which might lead one to hope that \ref{asmp: embedding} is
  non-essential. However, it is not possible to eliminate condition \ref{asmp:
    embedding} without incurring an additional factor of at least
  $(\rho/\sigma)^d$ in \eqref{eqn: mixing_time_upper_bound}, achieved, for
  instance, when $\Csig$ is a dumbbell-like set consisting of two balls of
  diameter $\rho$ linked by a cylinder of radius
  $\sigma$. \citet{abbasi-yadkori2016, abbasi-yadkori2016a}  
  develop theory regarding bi-Lipschitz deformations of convex sets, wherein it
  is observed that star-shaped sets as well as half-moon shapes of the type we
  consider in Section \ref{sec: experiments} both satisfy \ref{asmp: embedding}
  for reasonably small values of $L$. 
\end{remark}

@@

\begin{figure}
  \centering
%  \begin{adjustbox}{minipage=\linewidth}
%    \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.33\textwidth]{example1plots/sample2}
%      \caption{}
%    \end{subfigure}
%    \begin{subfigure}{.33\linewidth}
  \includegraphics[width=0.33\textwidth]{example1plots/sample1}
%      \caption{}
%    \end{subfigure}
%    \begin{subfigure}{.33\linewidth}
      \includegraphics[width=0.33\textwidth]{example1plots/sigma_normalized_cut_plot}
%      \caption{}
%    \end{subfigure}

%    \begin{subfigure}{.33\linewidth}
      \includegraphics[width=0.33\textwidth]{example1plots/sigma_mixing_time_plot}
%      \caption{}
%    \end{subfigure}
    \begin{subfigure}{.33\linewidth}
      \includegraphics[width=\linewidth]{example1plots/diameter_normalized_cut_plot}
      \caption{}
    \end{subfigure}
    \begin{subfigure}{.33\linewidth}
      \includegraphics[width=\linewidth]{example1plots/diameter_mixing_time_plot}
      \caption{}
    \end{subfigure}
    \caption{Samples, empirical results, and theoretical bounds for mixing time
      and normalized cut as diameter and thickness are varied. In (a) and (b),
      points in $\Cset$ are colored in red; points in $\Csig \setminus \Cset$
      are colored in yellow; and remaining points in blue.} 
    \label{fig:fig1}
  \end{adjustbox}
\end{figure}
